---
title: "SDS - project"
author: "Matteo Migliarini"
date: "`r Sys.Date()`"
output: html_document
---
```{r imports, include=FALSE}
library(tidyverse)
library(reshape2)
library(cowplot)
library(viridisLite)
library(LaplacesDemon)
data = read.csv('data/matteo-glasses-edited.csv') %>% 
  select(num_range('X', 1:11)) %>%
  replace(. == 'Yay', 1) %>%
  replace(. == 'Nay', 0) %>% 
  replace(. == '',   NA) %>%
  mutate_all(as.numeric) %>%
  as.matrix()
```

https://people.stat.sc.edu/hansont/stat740/jags_user_manual.pdf


# Data
The data consists in matrix $X$ of $45 \times 11$ observations, where each $X_{ij}$ is either an approval or a rejection of the $i$th rater to the $j$th pair of glasses (items).

Some of these datapoints are missing.

Below a graphical rappresentation of the dataset.
```{r values heatmap, echo=FALSE}
heatmap <- function(melted_data, x='x', y='y', value='value') melted_data %>%
  rename(x = x, y=y, value=value) %>%
  ggplot(aes(x,y, fill=value)) +
  geom_raster() +
  theme_void() + theme(aspect.ratio=1) + scale_fill_viridis_c(limits=c(0,1))

heatmap(melt(data), 'Var2', 'Var1')
```

# Jags
```{r}
library('R2jags')
jags_output_as_df <- function(jags_out) {
  bugs_out = jags_out$BUGSoutput$sims.array
  df = as.data.frame(bugs_out)
  colnames(df) = dimnames(bugs_out)[[3]]
  
  df
}

ITER = 10000
BURNIN = 1

I = nrow(data)
J = ncol(data)
jags_data = list(X = data, I = I, J = J)
```

## Dirichlet 
We model both the judges' attitude and the items' quality with a Dirichlet distribution.

```{r echo=FALSE}
system('cat models/dirichlet.txt')
```


```{r}
init.dir = list(init1 = list(psi=rep(1/J, J), tau=rep(1/I, I)))

jags.dir = jags(jags_data,
     model.file = 'models/dirichlet.txt',
     n.iter = ITER + BURNIN,
     n.burnin = BURNIN,
     inits = init.dir,
     n.thin = 1,
     n.chains = length(init.dir),
     parameters.to.save = c('psi', 'tau', 'theta'),
)

dir.mcmc = jags_output_as_df(jags.dir)
jags.dir$BUGSoutput$DIC

```


```{r}
get_theta_grid <- function(data) data %>%
  select(starts_with('theta')) %>%
  melt() %>%
  group_by(variable) %>%
  summarise(value = median(value)) %>%
  mutate(
    y=sub("theta\\[(\\d+),\\d+\\]", "\\1", variable) %>% as.numeric,
    x=sub("theta\\[\\d+,(\\d+)\\]", "\\1", variable) %>% as.numeric,
  )

dir.mcmc %>%
  get_theta_grid() %>%
  heatmap()
```

```{r}
jags.trace <- function(data, var) data %>%
  select(var) %>%
  melt() %>%
  rowid_to_column("index") %>%
  ggplot(aes(x=index,y = value)) +
  geom_line() #+ ylim(0,1)
  
```

La maggior parte delle variabili e' poco stabile
```{r}
# variables of which we want to see the trace plot
vars.tplot = c('theta[13,11]', 'theta[13,10]','tau[11]', 'tau[14]', 'psi[2]', 'psi[11]')
tplots.dir = vars.tplot %>%
  map(~ jags.trace(dir.mcmc, .))
names(tplots.dir) = vars.tplot 

plot_grid(plotlist = tplots.dir,
          ncol = 2, labels=vars.tplot)
```

SOME VALUES ARE OVER 1, THAT'S STRANGE
```{r}
t = dir.mcmc %>%
  select(starts_with('theta')) %>%
  melt() %>%
  mutate(
    x=sub("theta\\[(\\d+),\\d+\\]", "\\1", variable) %>% as.numeric,
    y=sub("theta\\[\\d+,(\\d+)\\]", "\\1", variable) %>% as.numeric,
  )

sum(t$value > 1)
```


```{r}
get_vector_variable <- function(data, var) data%>%
  select(starts_with(var)) %>%
  melt() %>%
  mutate(
    variable=sub('\\w+\\[(\\d+).*', '\\1', variable) %>% as.numeric
  )

dir.mcmc %>%
  get_vector_variable('psi') %>%
  ggplot(aes(x=variable, group=variable, y = value)) +
  geom_boxplot()
  
```
# Independent Betas
```{r}
system('cat models/independent-betas.txt')
```
```{r}
init.betas = list(init1 = list(psi=rep(0.5, J), tau=rep(0.5, I)))

jags.betas = jags(jags_data,
     model.file = 'models/independent-betas.txt',
     n.iter = ITER + BURNIN,
     n.burnin = BURNIN,
     inits = init.betas,
     n.thin = 1,
     n.chains = length(init.betas),
     parameters.to.save = c('psi', 'tau', 'theta'),
)

betas.mcmc = jags_output_as_df(jags.betas)

jags.betas$BUGSoutput$DIC
```

```{r}
betas.mcmc %>%
  get_theta_grid() %>%
  heatmap()
```


```{r}
tplots.betas = vars.tplot %>%
  map(~ jags.trace(betas.mcmc, .))
names(tplots.betas) = vars.tplot 

plot_grid(plotlist = tplots.betas,
          ncol = 2, labels=vars.tplot)
```

```{r}
betas.mcmc %>% 
  get_vector_variable('psi') %>%
  ggplot(aes(x=variable, group=variable, y = value)) +
  geom_boxplot()
```
# Multivariate Beta
```{r}
system('cat models/multivariate-beta.txt')
```
```{r}
generate_corr <- function(size, cor=0.2) {
  m = matrix(rep(cor,size^2), ncol=size)
  diag(m) <- 1
  m
}
```

```{r}
init.mvbeta = list(init1 = list(z_psi=rep(0, J), z_tau=rep(0, I)))
params.mvbeta = list(
  Sigma_psi = generate_corr(J,-0.01), 
  Sigma_tau = generate_corr(I,0))


jags.mvbeta = jags(c(jags_data, params.mvbeta),
     model.file = 'models/multivariate-beta.txt',
     n.iter = ITER + BURNIN,
     n.burnin = BURNIN,
     inits = init.mvbeta,
     n.thin = 1,
     n.chains = length(init.mvbeta),
     parameters.to.save = c('tau', 'psi', 'theta'),
)

mvbeta.mcmc = jags_output_as_df(jags.mvbeta)

jags.mvbeta$BUGSoutput$DIC
```
```{r}
mvbeta.mcmc %>%
  select(starts_with('tau'))

```


```{r}
mvbeta.mcmc %>%
  get_theta_grid() %>%
  heatmap()
```


```{r}
tplots.mvbeta = vars.tplot %>%
  map(~ jags.trace(mvbeta.mcmc, .))
names(tplots.mvbeta) = vars.tplot 

plot_grid(plotlist = tplots.mvbeta,
          ncol = 2, labels=vars.tplot)
```

```{r}
mvbeta.mcmc %>% 
  get_vector_variable('tau') %>%
  ggplot(aes(x=variable, group=variable, y = value)) +
  geom_boxplot()
```

# Item Response - Rasch
```{r}
system('cat models/rasch.txt')
```


```{r}
init.rasch = list(init1 = list(psi=rep(0, J), tau=rep(0, I)))

jags.rasch = jags(jags_data,
     model.file = 'models/rasch.txt',
     n.iter = ITER + BURNIN,
     n.burnin = BURNIN,
     inits = init.rasch,
     n.thin = 1,
     n.chains = length(init.rasch),
     parameters.to.save = c('tau', 'psi', 'theta'),
)

rasch.mcmc = jags_output_as_df(jags.rasch)

jags.rasch$BUGSoutput$DIC
```



```{r}
rasch.mcmc %>%
  get_theta_grid() %>%
  heatmap()
```


```{r}
tplots.rasch = vars.tplot %>%
  map(~ jags.trace(rasch.mcmc, .))
names(tplots.rasch) = vars.tplot 

plot_grid(plotlist = tplots.rasch,
          ncol = 2, labels=vars.tplot)
```

```{r}
rasch.mcmc %>% 
  get_vector_variable('tau') %>%
  ggplot(aes(x=variable, group=variable, y = value)) +
  geom_boxplot()
```


# Item Response - 2 parameters
```{r}
system('cat models/irt-2pl.txt')
```


```{r}
init.irt_2pl = list(init1 = list(psi=rep(0, J), tau=rep(0, I), alpha=rep(1,I)))

jags.irt_2pl = jags(jags_data,
     model.file = 'models/irt-2pl.txt',
     n.iter = ITER + BURNIN,
     n.burnin = BURNIN,
     inits = init.irt_2pl,
     n.thin = 1,
     n.chains = length(init.irt_2pl),
     parameters.to.save = c('tau', 'psi', 'theta', 'alpha'),
)

irt_2pl.mcmc = jags_output_as_df(jags.irt_2pl)

jags.irt_2pl$BUGSoutput$DIC
```



```{r}
irt_2pl.mcmc %>%
  get_theta_grid() %>%
  heatmap()
```
```{r}
jags.trace(irt_2pl.mcmc, 'alpha[7]')
```


```{r}
tplots.irt_2pl = vars.tplot %>%
  map(~ jags.trace(irt_2pl.mcmc, .))
names(tplots.irt_2pl) = vars.tplot 

plot_grid(plotlist = tplots.irt_2pl,
          ncol = 2, labels=vars.tplot)
```

```{r}
irt_2pl.mcmc %>% 
  get_vector_variable('alpha') %>%
  ggplot(aes(x=variable, group=variable, y = log(value))) +
  geom_boxplot()
```

# Item Response - Rasch
```{r}
system('cat models/rasch-hyerachical.txt')
```


```{r}
init.rasch_h = list(init1 = list(mu_psi = 0, sigma_psi = 1, tau=rep(0, I)))

jags.rasch_h = jags(jags_data,
     model.file = 'models/rasch-hyerachical.txt',
     n.iter = ITER + BURNIN,
     n.burnin = BURNIN,
     inits = init.rasch_h,
     n.thin = 1,
     n.chains = length(init.rasch_h),
     parameters.to.save = c('tau', 'psi', 'theta', 'mu_psi', 'sigma_psi'),
)

rasch_h.mcmc = jags_output_as_df(jags.rasch_h)

jags.rasch_h$BUGSoutput$DIC
```



```{r}
rasch_h.mcmc %>%
  get_theta_grid() %>%
  heatmap()
```


```{r}
tplots.rasch_h = vars.tplot %>%
  map(~ jags.trace(rasch_h.mcmc, .))
names(tplots.rasch_h) = vars.tplot 

plot_grid(plotlist = tplots.rasch_h,
          ncol = 2, labels=vars.tplot)
```

```{r}
rasch_h.mcmc %>% 
  get_vector_variable('psi') %>%
  ggplot(aes(x=variable, group=variable, y = value)) +
  geom_boxplot()
```
```{r}
jags.trace <- function(data, var, from=1, to=NULL) data %>%
  select(var) %>%
  melt() %>%
  rowid_to_column("index") %>%
  ggplot(aes(x=index,y = value)) +
  geom_line() + xlim(0,500)#+ ylim(0,1


jags.trace(rasch_h.mcmc, 'psi[1]')
```
```{r}
rasch_h.mcmc$mu_psi %>% sd
rasch_h.mcmc$sigma_psi %>% sd

```


# Altro